---
title: "Alaska bearded and spotted seal example dataset and analysis"
subtitle: "supporting material for the `crawl` package"
draft: false
author:
- name: Josh M. London
  affiliation: 1
- name: Devin S. Johson
  affiliation: 2
address:
- code: 1
  address: Alaska Fisheries Science Center, NOAA, Seattle, WA, USA 
  email: josh.london@noaa.gov
  orcid: orcid.org/0000-0002-3647-5046
- code: 2
  address: Alaska Fisheries Science Center, NOAA, Seattle, WA, USA 
  email: devin.johnson@noaa.gov
date: "`r format(Sys.Date(), '%d %B, %Y')`"
abstract: >
  The `crawl` package is an R package that provides analytical tools for
  estimating the movement paths of animals from telemetry data. The package is
  specifically designed with marine mammals in mind, but other species in other
  habits with similar telemetry devices will work. This package compendium
  provides step-by-step instructions for importing and setting up the telemetry
  data, initiating the `crwMLE()`, `crwPredict()`, and (eventually, when I get
  around to writing it up) `crwSimulator()` + `crwPostIS()` functions and
  visualizing results. The write up also discusses use of the `fix_path()`
  function to re-route predicted tracks around land. A few examples of data
  visualization plots are also presented. This compendium was originaly
  developed in support of a training course at the NOAA Fisheries Protected
  Species Assessment Workshop in La Jolla, California, USA.
output: 
  uswebr::html_uswds:
    number_sections: FALSE
---

## Introduction to Telemetry Data

The `crawl` package in R is designed, mostly, with telemetry data from 
satellite linked bio-loggers in mind. Locations derived from the ARGOS
satellite system or the GPS system (with data either recovered post-deployemnt
or transmitted during deployment) are the most common data source. In this
example, the data are all locations derived from the ARGOS satellite system.

The ARGOS system can have significant error associated with each location
estimate and those errors are described as either location qualities (3,2,1,0,A,B)
or, in more recent data, values that describe the ellipse error for each 
location. For this example, we will focus on the location quality values
and use a fairly generic set of values to describe each error quality.

The data for this example are available for download from the original source on
the [DataONE network](https://search.dataone.org/view/10.24431/rw1k118). These
were all deployments of bio-loggers from Wildlife Computers (Redmond,
Washington, USA) and each is a zipped archive pulled from the Wildlife Computers
Data Portal. Each archive contains both location and behavior data. For this
example, we will focus on just the location data.

please cite any use of the original data provided here with the following citation:

> Michael Cameron, Josh London, Kathy Frost, Alex Whiting, and Peter Boveng. Satellite Telemetry Dataset (Raw): Juvenile Bearded and Spotted Seals, 2004-2006, Kotzebue, Alaska. Research Workspace. 10.24431/rw1k118. 

And, if you are interested in learning more about the science, ecology, and
behavior of bearded (and spotted) seals, I recommend 
reading [this article](https://doi.org/10.1371/journal.pone.0192743) available
on PlosONE.

## Loading Raw Telemetry Data

All of the data are stored within the `data-raw` folder of our root
repository. We will use `list.files` and the `walk` function from
`purrr` to unzip everything into a temp directory.

```{r setup-load-data}
library(tidyverse)
library(here)
# here::here() provides a path relative to the project root directory
data_path <- here::here("data-raw")
td <- tempdir()

list.files(file.path(data_path), full.names=TRUE, pattern = "*.zip") %>%
  purrr::walk(~ unzip(.x, exdir = td, overwrite = TRUE)) 
```

The next step is to read just the _\*-Locations.csv_ file into R. The
`readr::read_csv()` function allows us to set the column data types
in advance so we don't have to mess with converting types afterwards. If you
have telemetry data from the Wildlife Computers data portal, these 
specifications should work for you as well.

```{r col-types}
my_cols <- cols(
  DeployID = col_character(),
  Ptt = col_integer(),
  Instr = col_character(),
  Date = col_datetime("%H:%M:%S %d-%b-%Y"),
  Type = col_character(),
  Quality = col_character(),
  Latitude = col_double(),
  Longitude = col_double(),
  `Error radius` = col_integer(),
  `Error Semi-major axis` = col_integer(),
  `Error Semi-minor axis` = col_integer(),
  `Error Ellipse orientation` = col_integer(),
  Offset = col_character(),
  `Offset orientation` = col_character(),
  `GPE MSD` = col_character(),
  `GPE U` = col_character(),
  Count = col_character(),
  Comment = col_character()
)
```

Now, we can combine the power of `readr::read_csv()` with `purrr:map_df()`
to read in all of the _\*-Locations.csv_ files into a single tibble (data.frame).

```{r read-csv}
tbl_locs <- list.files(file.path(td), full.names = TRUE, 
                       pattern = "*-Locations.csv") %>% 
  purrr::map_df( ~ readr::read_csv(.x, col_types = my_cols))

```

A key tennant of tidy data is that each row represents a single observation. For
location data, regardless of the data source, this is typical of the source data
structure. Each line in a file usually represents a single location estimate at
a specific time. So, there's very little we need to do in order to get our
`tbl_locs` into an acceptable structure.

One thing we can do, is to adjust the column names so there are no spaces or
hyphens. We can also drop everything to lower-case to avoid typos. To do this
we will rely on the `clean_names()` function within the `janitor` package. 
We also want to change the `date` column to a more appropriate and less 
confusing `date_time`. And, just because, I prefer `deployid` over `deploy_id`.

```{r}
library(janitor)
tbl_locs <- tbl_locs %>% 
  janitor::clean_names() %>%  
  dplyr::rename(date_time = date,
                deployid = deploy_id) %>% 
  dplyr::arrange(deployid, date_time)
```

Let's get a quick summary of our deployments

```{r deploy-summary}
tbl_locs %>% dplyr::group_by(deployid) %>% 
  dplyr::summarise(num_locs = n(),
            start_date = min(date_time),
            end_date = max(date_time))
```

There are some locations that were from test transmissions at the
Wildlife Computers office in Redmond, Washington, USA and at St. Andrews
in Scotland for a few SMRU tags. We can filter those out based on the
longitude values.

```{r remove-test-tx}
tbl_locs <- tbl_locs %>% 
  dplyr::filter(!between(longitude, -125, -120),
                !between(longitude, -50,0))
```

## Visualizing Source Data 

To further explore our imported data and confirm there aren’t any remaining 
issues to address, we can create visual representations of the data. The first
plot will rely on the spatial features within `ggplot2` and a custom `ptolemy`
package with world coastline data to create a plot of the `tbl_locs` object.
You can read up more on the `ptolemy` package on 
[the GitHub repo](https://github.com/jmlondon/ptolemy) and use
`devtools::install_github('jmlondon/ptolemy')` to install.

The `sf` package provides the bulk of our spatial data framework. This is not
intended to be a full `sf` tutorial, but, where possible, we will highlight a few
features and functions to explain the workflow. Users are encouraged to spend
significant time exploring and understanding the `sf` package through existing
documentation and several examples and blog posts available online. We also load
the `ggspatial` package for some convenience elements when plotting maps with
`ggplot`. If you are going to be making a lot of maps with `sf` and `ggplot2`,
I strongly suggest learning and adopting the functions within `ggspatial`. You
will save yourself a lot of time futzing with bounding boxes and other
parameters.

```{r spatial-libs}
library(sf)
library(ggspatial)

# library(devtools)
# devtools::install_github('jmlondon/ptolemy')
# after install, load and follow prompts to download source data
library(ptolemy)
```

Here, we are simply creating an `sf_locs` object via the `sf::st_as_sf()`
function. We specify the coordinate columns and then indicate the projection 
of the data is geographic (lat/lon) as indicated by the _4326_ epsg code.

In addition to points, it is often useful to organize the point data into
separate tracks for each deployment. The code to derive tracks from point data
within `sf` is a bit more complicated but is demonstrated below. It is important
that you include the `do_union = FALSE` argument to the `dplyr::summarise()`
function to maintain point order when creating the lines (and, in general, when
in doubt, insert a `dplyr::arrange(deployid, date_time)` call. The
`sf::st_cast("MULTILINESTRING")` step at the end converts the final geometry for
our object to _MULTILINESTRING_.

```{r create-sf-objects}
sf_locs <- sf::st_as_sf(tbl_locs, coords = c("longitude","latitude")) %>% 
  sf::st_set_crs(4326)

sf_lines <- sf_locs %>% 
  dplyr::arrange(deployid, date_time) %>% 
  dplyr::group_by(deployid) %>% 
  dplyr::summarise(do_union = FALSE) %>% 
  sf::st_cast("MULTILINESTRING")
```

The `sf_lines` object is, essentially, a data frame with each row representing 
a single line from a single deployment.

```{ r show-sf-lines}
sf_lines
```

These data are still in the _WGS84_ geographic coordinate system. Projecting the
data into a more appropriate coordinate reference system is strongly
encouraged and needed for some functionality in `crawl`. This is especially
important when/if data cross the 180 line, are spread over a large
latitudinal range, or within the polar regions.

```{r reproject}
sf_locs <- sf_locs %>% 
  sf::st_transform(3571)
sf_lines <- sf_lines %>% 
  sf::st_transform(3571)
```

The `ptolemy` package provides a number of pre-built collections of coastline
data as `sf` objects. In this case, we are interested in Alaska and the 
Bering Sea.

```{r ak-bering-coastline}
bering <- ptolemy::bering()
```

And, now, we build the `ggplot` using our `sf_lines` object. The
`annotation_spatial()` and `layer_spatial()` functions are helpers for 
handling some of the complexities for plotting spatial data with `ggplot2`. 
Most notably, it auto-picks a sensible bounding box.

```{r plot-raw-lines}
ggplot() + 
  annotation_spatial(bering, fill = "grey", lwd = 0) +
  layer_spatial(sf_lines, size = 0.75,aes(color = deployid)) +
  theme(legend.position = "none") +
  scale_color_viridis_d() +
  scale_x_continuous(breaks = seq(-180, 180, by = 5)) +
  ggtitle("Observed Argos Location Paths", 
          subtitle = "bearded and spotted seals, Bering Sea")
```

We can also plot the individual points themselves. However, if we're not
careful, the `ggplot` and `geom_sf()` process can be overwelmed and take
several minutes to plot (this will be fixed in upcomming updates to `ggplto2` and
`sf`). We have `r nrow(sf_locs)` _POINT_ records in our
`sf_locs` object. Instead of sending an object of many _POINT_ records to
`geom_sf` it is better to send fewer _MULTIPOINT_ recrods that, each, contain
many points. In this case, we can group our points into _MULTIPOINT_
features based on the `deployid` column.

```{r plot-raw-points}
ggplot() + 
  annotation_spatial(bering, fill = "grey", lwd = 0) +
  layer_spatial(data = sf_locs %>% 
            dplyr::group_by(deployid) %>% 
            dplyr::summarise(), 
          size = 0.75, aes(color = deployid)) +
  theme(legend.position = "none") +
  scale_color_viridis_d() +
  scale_x_continuous(breaks = seq(-180, 180, by = 5)) +
  ggtitle("Observed Argos Location Points", 
          subtitle = "bearded and spotted seals, Bering Sea")
```

## Duplicate Times

Argos data often contain near duplicate records. These are 
identified by location estimates with the same date-time but differing 
coordinate or error values. In theory, `crawl::crwMLE()` can handle these
situations, but we have found it is more reliable to fix these records. The 
first option for fixing the records would be to eliminate one of the duplicate
records. However, it is often not possible to reliably identify which record is
more appropriate to discard. For this reason, we advocate adjusting the 
date-time value for one of the records and increasing the value by 10 seconds. 
To facilitate this, we will rely on the `xts::make.time.unique()` function.

```{r dup-times, warning=FALSE, message = FALSE}
make_unique <- function(x) {
  suppressWarnings(
  xts::make.time.unique(x$date_time,eps = 10)
  )
}

library(xts)

tbl_locs <- tbl_locs %>% 
  dplyr::arrange(deployid,date_time) %>% 
  dplyr::group_by(deployid) %>% tidyr::nest() %>% 
  dplyr::mutate(unique_time = purrr::map(data, make_unique)) %>% 
  tidyr::unnest() %>% 
  dplyr::select(-date_time) %>% rename(date_time = unique_time)
```

## Course Speed Filter

This step is optional, but, as we see in this dataset, it is very common for
Argos data to include obviously wrong locations (locations that are many
kilometers away from the study area). Including such obviously wrong locations
in the analysis can result in unexpected issues or problems fitting. For this
reason, we recommend a course speed filter to remove these obviously wrong
locations. A typical speed filter might use a value of 2.5 m/s as a biologically
reasonable value for pinnipeds.

For this application, we will use 7.5 m/s (very conservative) and rely on the
`argosfilter` package. The example code below follows a typical
split-apply-combine approach using functions from the `dplyr`, `tidyr`, and
`purrr` packages. This analysis can be run in series, however, the process lends
itself nicely to parallel processing. The `parallel` package (included with the
distribution of R) would be one option for taking advantage of multiple
processors. However, we want to maintain the `purrr` approach so will use
`furrr` which brings the `future` parallel processing functionality to `purrr`.

First, lets print out the number or records in each deployment

```{r n-records}
tbl_locs %>% 
  dplyr::arrange(deployid, date_time) %>% 
  dplyr::group_by(deployid) %>% 
  tidyr::nest() %>% 
  mutate(n_records = map_int(data,nrow))
```

And, now, speed filter across multiple processors

```{r speed-filter, message=FALSE}
library(purrr)
library(furrr)

future::plan(multiprocess)

tbl_locs <- tbl_locs %>% 
  dplyr::arrange(deployid, date_time) %>% 
  dplyr::group_by(deployid) %>% 
  tidyr::nest() %>% 
  dplyr::mutate(filtered = furrr::future_map(data, ~ argosfilter::sdafilter(
    lat = .x$latitude,
    lon = .x$longitude,
    dtime = .x$date_time,
    lc = .x$quality,
    vmax = 7.5
  ))) %>% 
  tidyr::unnest() %>% 
  dplyr::filter(filtered %in% c("not", "end_location")) %>% 
  dplyr::select(-filtered) %>% 
  dplyr::arrange(deployid,date_time)
```

Let's see if there are any differences from our previous count of records

```{r n-records-post}
tbl_locs %>% 
  dplyr::arrange(deployid, date_time) %>% 
  dplyr::group_by(deployid) %>% 
  tidyr::nest() %>% 
  mutate(n_records = map_int(data,nrow))
```

Note, the previous two steps were done on our `tbl_locs` object which is
not an `sf` type and still in the geographic coordinate system. The
`argosfilter` package requires geopgraphic data and cannot accept an `sf`
object. So, we used `tbl_locs`. But, now, we need to convert and project.

```{r update-sf}
sf_locs <- sf::st_as_sf(tbl_locs, coords = c("longitude","latitude")) %>% 
  sf::st_set_crs(4326) %>% 
  sf::st_transform(3571)

sf_lines <- sf_locs %>% 
  dplyr::arrange(deployid, date_time) %>% 
  dplyr::group_by(deployid) %>% 
  dplyr::summarise(do_union = FALSE) %>% 
  sf::st_cast("MULTILINESTRING")
```

And, then, we can plot our data again

```{r plot-raw-lines-update}
ggplot() + 
  annotation_spatial(bering, fill = "grey", lwd = 0) +
  layer_spatial(sf_lines, size = 0.75,aes(color = deployid)) +
  theme(legend.position = "none") +
  scale_color_viridis_d() +
  scale_x_continuous(breaks = seq(-180, 180, by = 5)) +
  ggtitle("Observed Argos Location Paths (speed filtered)", 
          subtitle = "bearded and spotted seals, Bering Sea")
```

## Interactive Plots to Explore Tracks

The `mapview` package provides an easy means for creating interactive
maps of your spatial data. The one caveat to `mapview` is that it is based
on the Web Mercator projection (a la Google Maps, etc) and there are some
extra steps required if your tracks cross the dateline at 180 (which ours do ...
yay! fun!)

What we need to do is transform our data back to geographic and, then,
convert the coordinates from -180:180 to 0:360. We will use a custom
written function to handle this for us.

```{r st-to-360}
st_to_360 <- function(g) {
  coords <- (sf::st_geometry(g) + c(360,90)) %% c(360) - c(0,90)
  g <- sf::st_set_geometry(g,coords) %>% sf::st_set_crs(4326)
  return(g)
}
```

We will use the ESRI Ocean Basemap for our background layer. The default view
for `mapview` would create a legend with the colors for each `deployid` listed.
My preference is for something less cluttered. We want each `deployid` to show up
as a layer that can be turned on/off. To do this, we specificy `burst = TRUE` and
turn off the `legend` and `homebutton`.

```{r wrap-dateline}
library(mapview)
sf::st_transform(sf_lines,4326) %>% st_to_360() %>% 
mapview::mapview(map.types = "Esri.OceanBasemap",zcol = "deployid",
                 burst = TRUE, legend = FALSE, homebutton = FALSE)
```

## Create a Nested Tibble

When dealing with large numbers of deployments (and, potentially, wanting
to take advantage of parallel processing), book-keeping and data management 
are important.

We suggest organizing your data and analysis results into nested tibbles. This is
easily created with the `tidyr::nest()` function. In this case, we will organize
based on the `deployid` group.

```{r create-nested-tbl}
tbl_data <- sf_locs %>% 
  dplyr::group_by(deployid) %>% 
  nest()
tbl_data
```

## Fit Movement Model with `crawl::crwMLE()`

To fit a movement model to each collection of observed points, we will use
the `crawl::crwMLE()` function. The parameters below are a fairly generic
set of values. Researchers may need to customize these further for their
particular deployment and species characteristics.

The `fit_func()` is a wrapper that simplifies the downstream `purrr::map()`
function call.

Also, readers paying close attention may wonder if the `furrr` approach
described previously for running the speed filter would also work for
running `crwMLE`. The answer is, yes. But, we'll leave that for you to work
out on your own. I will advise that you read up on the `purrr::safely()`
function as a way to make sure a misfit deployment doesn't break the entire
process.

```{r fit-crwmle, message=FALSE}
# this doc and example requires the devel branch of `crawl` available
# on GitHub and installed with `devtools`
#
# library(devtools)
# devtools::install_github('NMML/crawl@devel)

library(crawl)
fixPar <- c(log(250), log(500), log(1500), rep(NA,3), rep(NA,2))

constr = list(lower=c(rep(log(1500),3), rep(-Inf,2)), upper=rep(Inf,5))

fit_func <- function(data,fixPar,constr) {
  suppressWarnings(
  crwMLE(mov.model = ~1, 
                   err.model=list(x=~quality-1), 
                   data=data, 
                   Time.name="date_time", 
                   fixPar=fixPar, 
                   constr=constr,
                   control=list(maxit=2000, trace=0, REPORT=1),
                   attempts=40,
                   initialSANN=list(maxit=200, trace=0, REPORT=1))
  )
}

# fit movement model
tbl_data <- tbl_data %>% 
  dplyr::mutate(fit = purrr::map(data,~fit_func(data = .x, constr = constr,
                                                fixPar = fixPar)))
```

## Predict Hourly Locations

Now that we have a model fit for each deployment, we want to predict hourly 
locations from fit movement model. This is payoff for all the work to this
point. By creating regular (in time) predictions for our deployments, we can
open the door to additional analysis (e.g. see the `momentuHMM` and `ctmcmove`
packages).

The time interval you choose for predicting is up to you. You can even provide
a vector of prediction times. For convenience, `predTime` accepts a character
string describing the time interval. In this case we use _1 hour_.

Lastly, you'll want to make sure the previous fit process was successful 
before predicting. The `crwMLE()` function should only return a `crwFit`
object if the fit was fully successful. That's what we are checking for in the
`filter` step below.

```{r predict-tracks, message = FALSE}

tbl_data <- tbl_data %>% 
  dplyr::filter(map_lgl(fit, ~inherits(.x,"crwFit"))) %>% 
  dplyr::mutate(pred_pts = purrr::map(fit, 
                                      ~crwPredict(.x, predTime = "1 hour")))
```

## Simulate Other Possible Tracks

The output from `crwPredict()` represents predictions along the _most likely
track_. But, we might also want to explore (or analyze) the broader range of
possible tracks. The combination of `crwSimulator()` and `crwPostIS` functions
can do this. 

> some point in the near future, I'll revisit this section and expand with
example code and plots for simulated tracks

## Convert predicted points to `sf` tracks

Now that we have our tracks fit and predicted paths created, we want to 
plot them and explore the results. 

The first thing we need to do is convert the `crwPredict` results to `sf`
objects. We can do that with a convenience function within the latest (remember,
you'll need the devel branch from GitHub) version of `crawl`. Note that we
specify `locType = "p"`. By default, `crwPredict` returns predictions at both
the specified interval (in this case, hourly) and the time point for each
corresponding observed location. By specifying `locType = "p"` we only get the
predicted locations at the regular interval of _1 hour_.

```{r crwpredict-as-sf}
tbl_data <- tbl_data %>% 
  dplyr::mutate(
    pts_sf = purrr::map(pred_pts, ~ crawl::crw_as_sf(.x, ftype = "POINT",
                                                     locType ="p")),
    line_sf = purrr::map(pred_pts, ~ crawl::crw_as_sf(.x, ftype = "LINESTRING",
                                                      locType = "p"))
    )
tbl_data
```

Lets _unnest_ our lines and plot them with `mapview`. 

```{r mapview-pred-lines, warning = FALSE}
tbl_data %>% 
  dplyr::select(deployid,line_sf) %>% 
  tidyr::unnest() %>%
  dplyr::select(-id) %>% 
  sf::st_as_sf(crs = 3571) %>% 
  sf::st_transform(4326) %>% st_to_360() %>% 
mapview::mapview(map.types = "Esri.OceanBasemap", zcol = "deployid",
                 burst = TRUE, legend = FALSE, homebutton = FALSE) 
```

## Fix predicted path around land

Bearded and spotted seals are coastal during some of their behavior and it isn't
unusual for the predicted tracks to cross land at times. 

```{r check-for-onland-pts, warning=FALSE, message=FALSE}
tbl_data %>% 
  dplyr::select(deployid, pts_sf) %>% 
  tidyr::unnest() %>%
  sf::st_as_sf(crs = 3571) %>% 
  sf::st_intersects(bering) %>% 
    purrr::map_lgl(~ length(.x) > 0) %>% 
  sum()
```

We, ideally, would include a land mask as part of the crwMLE model fit. But,
that becomes computationally challenging. As a compromise, `crawl` includes a
`fix_path()` function that (*is very much in beta development*) will re-route
portions of the predicted path around land.

```{r fix-path, warning=FALSE, message=FALSE}
tbl_data <- tbl_data %>%
  dplyr::mutate(fix = purrr::map2(
    pred_pts,
    fit,
    ~ crawl::fix_path(.x,
                      vector_mask = bering,
                      crwFit = .y)
  ))
```

Now, let's update our nested tibble with _\*_fix_ columns. We can also take a
moment to examine the number of points that may still intersect with our
land polygon (fingers crossed the resulting value is 0).

```{r fix-as-sf, warning=FALSE}
tbl_data <- tbl_data %>% 
  dplyr::mutate(
    pts_fix = purrr::map(fix, ~ crawl::crw_as_sf(.x, ftype = "POINT",
                                                      locType = "p")),
    line_fix = purrr::map(fix, ~ crawl::crw_as_sf(.x, ftype = "LINESTRING",
                                                      locType = "p"))
    )

tbl_data %>% 
  dplyr::select(deployid,pts_fix) %>% 
  tidyr::unnest() %>%
  sf::st_as_sf(crs = 3571) %>% 
  sf::st_intersects(bering) %>% 
    purrr::map_lgl(~ length(.x) > 0) %>% 
  sum()
```

Now, let's go ahead and re-create our `mapview` and zoom in closely to some
of the coastline and see how things look. One thing to keep in mind is that
the tracklines may still cross land at some locations since we only predicted
every hour. There's a tradeoff: the more frequent predictions, the smoother
the path and the fixed path around land features. But, the computational
effort will increase.

If you want to challenge your skills, go ahead and create a `mapview` that plots
`pts_fix` instead of `line_fix`. But, remember the cautionary bit earlier about
creating _MULTIPOINT_ features instead of sending large _POINT_ features to the
map.

```{r, warning=FALSE}
tbl_data %>% 
  dplyr::select(deployid,line_fix) %>% 
  tidyr::unnest() %>%
  dplyr::select(-id) %>% 
  sf::st_as_sf(crs = 3571) %>% 
  sf::st_transform(4326) %>% st_to_360() %>% 
mapview::mapview(map.types = "Esri.OceanBasemap", zcol = "deployid",
                 burst = TRUE, legend = FALSE, homebutton = FALSE) 
```

## Animated Plot

Now, let's make something really cool and animate the tracks. We'll use the
`gganimate` package. This will probably take about 7-10 minutes to render.

```{r animate, warning=FALSE, message=FALSE, fig.height=8}
library(gganimate)
library(ggspatial)


pts_by_day <- tbl_data %>% 
  dplyr::select(deployid,pts_fix) %>% 
  tidyr::unnest() %>%
  dplyr::filter(locType == "p") %>% 
  dplyr::select(deployid,date_time, geometry) %>% 
  dplyr::mutate(anim_date = format(date_time, format="%B %d"),
                yday = lubridate::yday(date_time)) %>% 
  sf::st_as_sf(crs = 3571) %>% 
  dplyr::arrange(deployid,yday) %>% 
  dplyr::group_by(deployid,yday) %>% 
  dplyr::summarise(do_union = FALSE)

animated_plot <- ggplot() +
  annotation_spatial(bering, fill = "grey", lwd = 0) +
  layer_spatial(pts_by_day, size = 0.75,aes(color = deployid)) +
  scale_color_viridis_d() + 
  scale_x_continuous(breaks = seq(-180, 180, by = 5)) +
  theme(legend.position = "none") +
  ggtitle("Predicted Movements, Day of Year: {as.integer(frame_time)}", 
          subtitle = "bearded and spotted seals, Bering Sea") +
  transition_time(yday) +
  shadow_wake(wake_length = 0.1, alpha = FALSE)

gganimate::animate(animated_plot)
```
